---
title: "Covid91 vaccine study Final2023"
author: "Estevan Green"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$ : The covid vaccine  is not  effective in treating covid among subgroups.


$H_A$ : The covid vaccine is effective in treating  covid among subgroups.


## Methods 


The statistical analysis well be using is cat~ cat because the covid vaccine should be effective among subgroups.

The method we will use to display the data is a bar chart,cross tabel , and chi squared test 

## Males


## Results
```{r}
maleData = filter (FinalData,sex == "M")
```

## Graphical Results 
```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")

```

## Numerical Graphical Results 

```{r}
table1 <- xtabs(~infected + treatment , data=maleData)
rowPerc(table1)
colPerc(table1)
```

## Inferental Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

```{r}
femaleData = filter (FinalData,sex == "F")
```

## Females

## Graphical Results 
```{r}
barchartGC(~infected+treatment,data=femaleData)
barchartGC(~infected+treatment,data=femaleData, type="percent")

```

## Numerical Graphical Results 

```{r}
table2 <- xtabs(~infected + treatment , data=femaleData)
rowPerc(table2)
colPerc(table2)
```
## Inferental Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```


## LGBTQ

```{r}
LGBTQData = filter (FinalData,LGBTQ == "gay")
```

## Graphical Results 

```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")

```

## Numerical Graphical Results 
```{r}
table3 <- xtabs(~infected + treatment , data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

## Inferental Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```




## Druggies

```{r}
DruggiesData = filter (FinalData, DrugUser == "yes")
```
## Graphical Results 

```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")

```
## Numerical Graphical Results 

```{r}
table4 <- xtabs(~infected + treatment , data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```

## Inferental Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```


# Overall Results and Conclusions