---
title: "Covid91 vaccine study Final2023"
author: "Estevan Green"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---
```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$ The COVID vaccine has no effect on people in different subgroups.


$H_A$ The COVID vaccine does have an effect on people different subgroups.

## Methods 

This statistical analysis we will be using is catergorical ~ categorical, because the COVID vaccine should be effective among subgroups

The method we will be using to display the data us a bar chart, cross table, and chisq test

# Males

## Results

```{r}
maleData = filter (FinalData,sex == "M")
```
## Descriptive Results

### Graphical Descriptive 

```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")

```

Based on the graph, it is showing us that the group of people who contracted COVID and the ones who received the vaccine had less symptoms than those that received the placebo, while the all male group that didn't contract COVID showed no difference between those who were vaccinated and unvaccinated. 

### Numerical Descriptive

```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

This table is showing that 69.57% of the unvaccinated male group experienced symptoms while only 30.43% of the vaccinated males experienced symptoms. It is also telling us that the male placebo group, 2.55% of them contracted COVID while only 1.15% of the vaccinated group contracted COVID.

## Descriptive Results

### Graphical Descriptive 

```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")

```

Based on the graph, it is showing us that the group of people who contracted COVID and the ones who received the vaccine had less symptoms than those that received the placebo, while the all male group that didn't contract COVID showed no difference between those who were vaccinated and unvaccinated. 

### Numerical Descriptive

```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

This table is showing that 69.57% of the unvaccinated male group experienced symptoms while only 30.43% of the vaccinated males experienced symptoms. It is also telling us that the male placebo group, 2.55% of them contracted COVID while only 1.15% of the vaccinated group contracted COVID.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value is 0.0000000000001098, concluding that the p value is so small that we are rejecting the null hypothesis. Showing that COVID vaccine is effective between the male subgroup.

Based on the Fisher exact test, we got a p-value of 0.0000000000004631 concluding this is giving a strong basis for the observed association between variables meaning we are rejecting the null hypothesis. The odds in this fisher test are showing males have a 2.24% chance of the events odds happening in one group gicing more evidence to the positive association of the variables.

# Females

## Results

```{r}
femaleData = filter (FinalData,sex == "F")
```

## Descriptive Results

### Graphical Descriptive 

```{r}
barchartGC(~infected+treatment,data=femaleData)
barchartGC(~infected+treatment,data=femaleData, type="percent")
```

This graph is showing that from the all female group who contracted COVID, the ones who received the vaccine had less symptoms than those with the placebo while the all the female group that did not contract COVID showed no difference between those who were vaccinated and not vaccinated. 

### Numerical Descriptive

```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```
The data from this table shows that from the female group that were infected with COVID 19, those who were unvaccinated, 60% of them experienced symptoms while only 40% of the female group who were vaccinated experienced symptoms. It is also showing that within the female placebo group, 2.04% were infected with COVID, while only 1.37% of the vaccinated group contracted COVID.

### Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

Based on this chisq test, the p-value is 0.0002732, showing that there is a significant difference, concluding that we are rejecting the null hypothesis. This means that the COVID vaccine is effective within the female subgroup. 

For the fisher exact test show that the odds are about 1.50%, suggesting that there is a significant correlation between the variables, showing that there is strong evidence directing that we are rejecting the null hypothesis. 

